---
title: 'Arizona Groundwater research'
author: "[Angus Watters](https://anguswg-ucsb.github.io/)"
output:
  mikedown::ucsb:
    toc: TRUE
    number_sections: FALSE
    code_folding: "hide"
---


```{r, message= FALSE, warning = FALSE, echo = FALSE}

library(tidyverse)
library(rmarkdown)     # You need this library to run this template.
library(mikedown)   
```



***

```{r, message= FALSE, warning = FALSE, echo = TRUE}
# Libraries
library(tidyverse)
library(sf)
library(USAboundaries)
library(sp)
library(leaflet)
library(plotly)
library(lubridate)
library(RColorBrewer)
```

```{r, message= FALSE, warning = FALSE, echo = FALSE}
source("../docs/utils.R") 
```

<br>
<br>

# **Measurement threshold** *(1920 - 2019)*
A minimum threshold of 10 distinct measurements with at least 1 measurement every 5 years was applied to filter for only regularly monitored wells 

***

<br>
<br>


```{r, message= FALSE, warning = FALSE, echo = FALSE}
join_time = readRDS("../data/az/az-join-time.rds")
join_spatial = readRDS('../data/az/az-join-spatial.rds')

# THRESHOLDING 

join_spatial = join_spatial[!duplicated(join_spatial$wellid),]

join_time = join_time %>% filter(year_dist >= 5)
join_spatial = join_spatial %>% filter(wellid %in% join_time$wellid)

join_time = join_time %>% 
  group_by(wellid)
join_spatial = join_spatial %>% 
  group_by(wellid)
```

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
font = list(
    family = 'Courier',
    size = 15,
    color = 'white'
  )
label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font
  )

x = join_spatial %>%
  st_transform(4326) %>% 
  group_by(wellid)

# Arizona state shape (CRS = 5070)
az = us_states() %>%
  filter(state_name == 'Arizona') %>% 
  st_transform(5070)


az_outline = az %>% 
  st_cast('MULTILINESTRING') %>% 
  st_transform(4326)


ama = read_sf('../data/az-ama-shps/Act_Man_Areas.shp') %>% 
  st_transform(5070) %>% 
  st_cast("MULTIPOLYGON")

ama2 = ama %>% st_transform(4326)

### AQUIFERS
aquifer = read_sf('../data/shps/us_aquifers.shp') %>% 
  st_transform(5070) %>% 
  st_cast("MULTIPOLYGON")

aquifer2 = aquifer %>% st_transform(4326)

box = az %>% st_transform(4326) %>% 
  st_bbox() %>% 
  st_as_sfc()
temp = st_intersects(aquifer2, box)

aquifer2 = aquifer2[which(lengths(temp) != 0), ]

# COLOR PALLETES 
#RColorBrewer::display.brewer.all(n=4, exact.n=FALSE)

nb.cols = 10
col1 = RColorBrewer::brewer.pal(9,"Blues")
# col2 = brewer.pal(9,"YlOrRd")
col2 = colorRampPalette(brewer.pal(9,"YlOrRd"))(nb.cols)
col3 = RColorBrewer::brewer.pal(9,"YlGnBu")
col4 = RColorBrewer::brewer.pal(9,"Spectral")
col5 = RColorBrewer::brewer.pal(9,"Greys")
col6 = palette(c('black', 'white'))

pals1 = colorFactor('cyan', domain = x$dtw)
pals2 = colorNumeric(col2, domain = c(0, 1500))
pals3 = colorBin(col3, domain = 1:8)
pals4 = colorFactor(col4, domain = aquifer2$AQ_NAME)
pals5 = colorFactor(col6, domain = x$source)

# LEGEND LABELS
labels = c('0', '100', '200',
           '300', '400',
           '500', '600',
           '700', '800', '900')
agency_labs = c('ADWR', 'USGS')
# records every ~6 months
adwr_spatial = x %>% 
  filter(source == 'AZ') %>% 
  select(wellid, dtw, date, source, measurement_dist)
adwr_spatial = adwr_spatial %>%
  # filter(dtw < 1000, dtw >= 0) %>% 
  select(wellid, dtw, date, source, measurement_dist)

# records every ~ 1 month
usgs_spatial = x %>% 
  filter(source == 'USGS')
usgs_spatial = usgs_spatial %>%
  # filter(dtw < 1000, dtw >= 0) %>% 
  select(wellid, dtw, date, source, measurement_dist)

r0 =  dtw_range(x, -200, 0) %>% select(wellid, dtw, date,source, measurement_dist)


leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron, group = 'Base') %>% 
  addProviderTiles(providers$Esri.WorldTopoMap, group = "Terrain") %>% 
  addPolygons(data = ama2,
              fillColor  = ~pals3(OBJECTID),
              fillOpacity = 0.5,
              color = 'black',
              weight = 1,
              label = ~MAP_LABEL, group = 'AMA') %>%
  addPolygons(data = aquifer2,
              fillColor = ~pals4(AQ_NAME),
              fillOpacity = 0.3,
              color = 'black',
              weight = 1,
              label = ~AQ_NAME, group = 'Aquifer') %>%
  addPolygons(data = az_outline,
            fillOpacity = 0.1,
            color = 'black',
            weight = 2) %>%
  addCircleMarkers(data = usgs_spatial, #clusterOptions = markerClusterOptions(interactive()),
                   radius = 4,
                   fillColor = ~pals2(dtw), 
                   fillOpacity = .8, 
                   color = 'black',
                   # color = ifelse(r1$source == 'USGS', 'black', NA),
                   opacity = .8, 
                   weight = .5,
                   stroke = T,
                   popup = leafpop::popupTable(usgs_spatial, feature.id = FALSE,
                                               row.numbers = FALSE), group = 'USGS') %>% 
  addCircleMarkers(data = adwr_spatial,
                   radius = 4, #clusterOptions = markerClusterOptions(interactive()),
                   fillColor = ~pals2(dtw), 
                   fillOpacity = .8, 
                   color = 'black',
                   # color = ifelse(r2$source == 'USGS', 'black', NA),
                   opacity = .8, 
                   weight = .5,
                   stroke = T,
                   popup = leafpop::popupTable(adwr_spatial, feature.id = FALSE,
                                               row.numbers = FALSE), group = 'ADWR') %>% 
 addCircleMarkers(data = r0, #clusterOptions = markerClusterOptions(interactive()),
                  radius = 4,
                   fillColor = ~pals1(dtw), 
                   fillOpacity = .8, 
                  color = 'black',
                   #color = ifelse(r0$source == 'USGS', 'black', NA),
                   opacity = .8,
                   weight = .5,
                   stroke = TRUE,
                   popup = leafpop::popupTable(r0, feature.id = FALSE,
                                               row.numbers = FALSE), group = 'Negative depths') %>% 
  addLayersControl(overlayGroups = c('USGS', 'ADWR', 'Negative depths', 'AMA', 'Aquifer'),
                   options = layersControlOptions(collapsed = FALSE),
                   baseGroups = c("Base", 'Terrain')) %>% 
  addLegend(pal = pals2,
            values = c(0, 1500),
            opacity = .9,
            title = 'Depth to water (ft)', # Title
            position = "bottomleft",
            labFormat = function(type, cuts, p) {
              paste0(labels)}) 

```





***

<br>
<br>
<br>
<br>

<br>

# **Active Management Areas**
The 1980 Arizona Groundwater code established 8 Active management areas (AMA) in areas heavily reliant on mined groundwater. The goal of the AMAs is too achieve safe yield groundwater usage in these heavy groundwater pumping areas.

---

<br>
<br>
<br>

## **Joseph City AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Joseph City AMA
joseph_shp = ama2 %>% 
  filter(OBJECTID == 1) %>% 
  st_intersection(x) %>% 
  mutate(name = 'JOSEPH CITY')
joseph_wells = join_time %>% filter(wellid %in% joseph_shp$wellid)


plot_test1 = ggplot(data = joseph_wells, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

joseph_ggplotly = ggplotly(plot_test1, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
joseph_ggplotly
```

<br>
<br>
<br>

## **Prescott AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Prescott AMA
presc_shp = ama2 %>% 
  filter(OBJECTID == 2) %>% 
  st_intersection(x) %>% 
  mutate(name = 'PRESCOTT')
presc_wells = join_time %>% filter(wellid %in% presc_shp$wellid)

plot_test2 = ggplot(data = presc_wells, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

prescott_ggplotly = ggplotly(plot_test2, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
prescott_ggplotly
```

<br>
<br>
<br>

## **Pheonix AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Pheonix AMA
pheonix_shp = ama2 %>% 
  filter(OBJECTID == 3) %>% 
  st_intersection(x) %>% 
  mutate(name = 'PHOENIX')
pheonix_wells = join_time %>% filter(wellid %in% pheonix_shp$wellid)

pheonix_filter = pheonix_wells %>% 
  group_by(wellid) %>%  filter(measurement_dist > 70)

plot_test3 = ggplot(data = pheonix_filter, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

ggplotly(plot_test3, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
```

<br>
<br>
<br>

## **Harquahala AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Harquahala AMA
harq_shp = ama2 %>% 
  filter(OBJECTID == 4) %>% 
  st_intersection(x) %>% 
  mutate(name = 'HARQUAHALA')
harq_wells = join_time %>% filter(wellid %in% harq_shp$wellid)

plot_test4 = ggplot(data = harq_wells, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

harq_ggplotly = ggplotly(plot_test4, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
harq_ggplotly
```

<br>
<br>
<br>

## **Pinal AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}

# Pinal AMA
pinal_shp = ama2 %>% 
  filter(OBJECTID == 5) %>% 
  st_intersection(x) %>% 
  mutate(name = 'PINAL')
pinal_wells = join_time %>% filter(wellid %in% pinal_shp$wellid)

plot_test5 = ggplot(data = pinal_wells, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

pinal_ggplotly = ggplotly(plot_test5, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
pinal_ggplotly
```

<br>
<br>
<br>

## **Tucson AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Tuscon AMA
tucson_shp = ama2 %>% 
  filter(OBJECTID == 6) %>% 
  st_intersection(x) %>% 
  mutate(name = 'TUCSON')
tucson_wells = join_time %>% filter(wellid %in% tucson_shp$wellid)
tucson_wells = tucson_wells %>% group_by(wellid)

tucson_filter =tucson_wells %>% 
  filter(wellid != 6008) %>% 
  mutate(time_span = max(year) - min(year)) %>% 
  filter(time_span > 50)

plot_test6 = ggplot(data = tucson_filter, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))


tucson_ggplotly = ggplotly(plot_test6, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
tucson_ggplotly
```

<br>
<br>
<br>

## **Douglas AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Douglas AMA
douglas_shp = ama2 %>% 
  filter(OBJECTID == 7) %>% 
  st_intersection(x) %>% 
  mutate(name = 'DOUGLAS')
douglas_wells = join_time %>% filter(wellid %in% douglas_shp$wellid)

plot_test7 = ggplot(data = douglas_wells, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

douglas_ggplotly = ggplotly(plot_test7, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)
douglas_ggplotly
```

<br>
<br>
<br>

## **Santa Cruz AMA**

---

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
# Santa Cruz AMA
santacruz_shp = ama2 %>% 
  filter(OBJECTID == 8) %>% 
  st_intersection(x) %>% 
  mutate(name = 'SANTA CRUZ')
santacruz_wells = join_time %>% filter(wellid %in% santacruz_shp$wellid)

plot_test8 = ggplot(data = santacruz_wells, aes(x = date, y = dtw)) +
  geom_line(aes(col = wellid), size = 1) +
  ylim(1000, 0) +
  geom_smooth(col = 'black') +
  labs(x = 'Year',
       y = 'DTW (ft)') +
  theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
        axis.text.x = element_text(color="black", size=14), 
        axis.text.y = element_text(color="black", size=14), 
        axis.title.x = element_text(face="bold", color="black", size=16), 
        axis.title.y = element_text(face="bold", color="black", size=16), 
        panel.grid.major = element_line(colour = "#808080"),
        panel.grid.minor = element_line(colour = "#808080", size = 1),
        strip.text.x = element_text(face = 'bold', size = 10))

santacruz_ggplotly = ggplotly(plot_test8, tooltip = c('x', 'y', 'wellid')) %>%
  style(hoverlabel = label) %>% 
  layout(font = font, 
         yaxis = list(fixedrange = TRUE)) %>% 
  config(displayModeBar = FALSE)

santacruz_ggplotly
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>


# **Buffer analysis on deep water wells**

***

Investigation of wells below 300ft in which depth has continually decreased. The first plot shows all wells within a certain depth range that have had continuous measurement records for over 50 years. The following plots focus on a specific well in that range, and then a buffer was formed around these wells (12 - 20 km) to then see what is happening to other wells within the same area. The buffer analysis was used on individual wells that displayed dramatic and continual decreases in depth to water.  


<br>
<br>
<br>




<br>

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
cont_records = join_time %>%
  filter(wellid != 6008) %>% 
  mutate(time_span = max(year) - min(year)) %>% 
  filter(time_span > 50)

```

## **Range: 300 - 500 ft**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
cont2 = cont_records %>% filter(time_span > 65)
plotRange(cont2, 300, 500)
```

***

<br>
<br>
<br>

### **Well 4573**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE,, out.width = "100%"}
# RETURNS A TIME SERIES PLOT OF WELLS WITHIN A BUFFER DISTANCE (km)
plotBuffer(join_spatial, 4573, 12500)
```

```{r, message= FALSE, warning = FALSE, echo=FALSE,, out.width = "100%"}
test = join_spatial %>% filter(wellid %in% c('4573', '4740', '3472', '4401',
                                             '4182', '4993', '5730', '4953', '1853', '5743'))

f = ggplot() +
  geom_sf(data = az) +
  geom_sf(data = test, aes(col = wellid), size = 2)
ggplotly(f)
```

<br>
<br>
<br>


### **Well 4740**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 4740, 12500)
```


<br>
<br>
<br>

### **Well 3472**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 3472, 12500)
```



<br>
<br>
<br>

## **Range: 400 - 500 ft**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotRange(cont_records, 400, 500)
```


<br>
<br>
<br>

### **Well 4401**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 4401, 12500)
```


<br>
<br>
<br>


### **Well 4182**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 4182, 12500)
```


<br>
<br>
<br>

## **Range: 500 - 600 ft** 

***


```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotRange(cont_records, 500, 600)
```


<br>
<br>
<br>

### **Well 4993**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 4993, 20000)
```


<br>
<br>
<br>

### **Well 5730**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 5730, 30000)
```

<br>
<br>
<br>

### **Well 4953**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 4953, 12500)
```

<br>
<br>
<br>

### **Well 1853**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 1853, 20000)
```

<br>
<br>
<br>

### **Well 5743**

***

```{r, message= FALSE, warning = FALSE, echo=FALSE, out.width = "100%"}
plotBuffer(join_spatial, 5743, 20000)
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>

# **Negative DTW measurements**

***

A negative DTW reading may indicate groundwater emerging above the surface in the form of a spring. There are 9 wells that recorded negative depth to water readings. In the area of Chino Valley there are 6 wells with negative DTW to water readings all within a 5 km radius of one another. They other 3 wells are located in the far southeast corner of Arizona, nearing the US-Mexico border.

## **Negative Case Rules**

***

<br>


- **Case 1a:** Artesian to non-artesian (negative to positive sequentially)
  - maintained these values

- **Case 1b:** Non-artesian to artesian (positive to negative sequentially, skeptical of these)
  - kept unless otherwise noted 
  
- **Case 1c:** Well with very low depth to water and fluctuation between slightly below ground surface and artesian 
  - maintained these values

- **Case 1d:** Well appears to have a pause in pumping activity (multiple years), returns to artesian, and then is pumped again 
  - maintained these values 

- **Case 2a:** Measurements are sequential except for a random negative, therefore was likely a recording error
  - changed negative value to positive

- **Case 2b:** Measurements are sequential except for periods of negative values that appear to be recorded incorrectly 
  - changed negative value to positive

- **Case 2c:** Measurements are sequential and within reason for well depth however all measurements for that well are negative and are likely recorded incorrectly
  - compared the well spatially with others to determine if artesian levels would be appropriate, and if not, changed to positive values

- **Case 3a:** Measurements are sequential except for truly wonky number which doesn’t make sense
  - removed this observation
  
- **Case 3b:** Measurements are sequential and then dataset changes to record 0 values only, or some other illogical number continuously, which does not make sense
  - removed these wells entirely 

***

<br>
<br>
<br>
<br>


```{r, message= FALSE, warning = FALSE, echo = FALSE}
cases = data.frame(cases = c('1a', '1b', '1c', '1d', '2a', '2b', '2c', '3a', '3b'), wells = c(3, 0, 6, 0, 0, 0, 3, 0, 1))

knitr::kable(cases, caption = 'Negative wells in Arizona',
             col.names = c('Cases', 'Occurances'))
```
### **Negative DTW plots**


```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
negatives = join_time %>% filter(year_dist >= 5, dtw <= 0)
neg_time = join_time %>% filter(wellid %in% negatives$wellid)
neg_time = neg_time %>% 
  mutate(case = case_when(wellid == 5490 ~ "1A",
                          wellid == 5534 ~ "1A",
                          wellid == 5538 ~ "1A",
                          wellid == 222 ~ "1C",
                          wellid == 226 ~ "1C",
                          wellid == 259 ~ "1C",
                          wellid == 2633 ~ "1C",
                          wellid == 5240 ~ "1C",
                          wellid == 5985 ~ "1C",
                          wellid == 1 ~ "2C",
                          wellid == 524 ~ "2C",
                          wellid == 5503 ~ "2C",
                          wellid == 5506 ~ "3B")) 


case_1a = neg_time %>% filter(case == '1A')
spatial_1a = join_spatial %>% filter(wellid %in% neg_time$wellid) %>% 
    mutate(case = case_when(wellid == 5490 ~ "1A",
                          wellid == 5534 ~ "1A",
                          wellid == 5538 ~ "1A",
                          wellid == 222 ~ "1C",
                          wellid == 226 ~ "1C",
                          wellid == 259 ~ "1C",
                          wellid == 2633 ~ "1C",
                          wellid == 5240 ~ "1C",
                          wellid == 5985 ~ "1C",
                          wellid == 1 ~ "2C",
                          wellid == 524 ~ "2C",
                          wellid == 5503 ~ "2C",
                          wellid == 5506 ~ "3B"))

case_1c = neg_time %>% filter(case == '1C')
case_2c = neg_time %>% filter(case == '2C')
case_3b = neg_time %>% filter(case == '3B')

map_1a = ggplot() +
  geom_sf(data = az) +
  geom_sf(data = spatial_1a, aes(fill = wellid, col = case), size = 2) +
  labs(fill = '',
    col = '(WELL, CASE)')

maply_1a = ggplotly(map_1a, tooltip = c('wellid', 'case')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font)
maply_1a


case_1c = neg_time %>% filter(case == '1C')
case_2c = neg_time %>% filter(case == '2C')
case_3b = neg_time %>% filter(case == '3B')

```

***

<br>
<br>
<br>
<br>
<br>



### **Cases**

***

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}

case_1c = neg_time %>% filter(case == '1C')
case_2c = neg_time %>% filter(case == '2C')
case_3b = neg_time %>% filter(case == '3B')
  font = list(
    family = 'Courier',
    size = 15,
    color = 'white')
  label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font)
  gg_1a = ggplot(data = case_1a, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    geom_hline(aes(yintercept = 0), size = 1) +
    scale_y_reverse() +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well',
         title = 'Case 1A') + 
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1),
          legend.title = element_text(colour="black", size=16, face="bold"),
          legend.text = element_text(colour="black", size=10, face="bold"))
  plot_1a = ggplotly(gg_1a, tooltip = c('x', 'y', 'wellid', 'case')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot_1a
```

<br>
<br>
<br>


***

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
    gg_1c = ggplot(data = case_1c, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    geom_hline(aes(yintercept = 0), size = 1) +
    scale_y_reverse() +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well',
         title = 'Case 1C') + 
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1),
          legend.title = element_text(colour="black", size=16, face="bold"),
          legend.text = element_text(colour="black", size=10, face="bold"))
  plot_1c = ggplotly(gg_1c, tooltip = c('x', 'y', 'wellid', 'case')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot_1c
```

<br>
<br>
<br>



***

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
  
    gg_2c = ggplot(data = case_2c, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    geom_hline(aes(yintercept = 0), size = 1) +
    scale_y_reverse() +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well',
         title = 'Case 2C') + 
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1),
          legend.title = element_text(colour="black", size=16, face="bold"),
          legend.text = element_text(colour="black", size=10, face="bold"))
  plot_2c = ggplotly(gg_2c, tooltip = c('x', 'y', 'wellid', 'case')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot_2c
```

<br>
<br>
<br>



***

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}
  
    gg_3b = ggplot(data = case_3b, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    geom_hline(aes(yintercept = 0), size = 1) +
    scale_y_reverse() +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well',
         title = 'Case 3B') + 
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1),
          legend.title = element_text(colour="black", size=16, face="bold"),
          legend.text = element_text(colour="black", size=10, face="bold"))
  plot_3b = ggplotly(gg_3b, tooltip = c('x', 'y', 'wellid', 'case')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot_3b
```



<br>
<br>
<br>


### **Chino Valley**

***

```{r, message= FALSE, warning = FALSE, echo = FALSE, out.width = "100%"}

neg4 = join_spatial %>% filter(wellid == 5534)
buff4 = st_buffer(join_spatial[neg4, ], 12500)
near_neg4 = st_intersection(join_spatial, buff4)
near_time4 = join_time %>% filter(wellid %in% near_neg4$wellid)
plotNegativeWells(near_time4)
```

<br>
<br>
<br>
<br><br>
<br>
<br>
<br>
<br>
<br>
<br>

# **Functions**

***


```{r, message= FALSE, warning = FALSE, eval=FALSE, echo=TRUE}
### PLOT AN INDIVIDUAL WELL BY WELL ID (DTW TIME SERIES)

### PLOT AN INDIVIDUAL WELL BY WELL ID (DTW TIME SERIES)
plotWell = function(df_time, num) {
  font = list(
    family = 'Courier',
    size = 15,
    color = 'white')
  label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font)
  well = df_time %>% filter(wellid == num)
  gg = ggplot(data = well, aes(x = date, y = dtw)) +
    geom_line(data = well, aes(y = dtw, col = wellid), size = 1) +
    ylim(max(well$dtw) + 100, 0) +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well') +  
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1))
  plot = ggplotly(gg, tooltip = c('x', 'y', 'wellid')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot
}

# PLOTS WELL DTW TIME SERIES WITHIN A RANGE (MIN, MAX)
plotRange = function(df_time, min, max) {
  font = list(
    family = 'Courier',
    size = 15,
    color = 'white')
  label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font)
  df_time = df_time %>% filter(dtw <= max, dtw >= min) %>% 
    arrange(desc(date))
  gg = ggplot(data = df_time, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    ylim(max(df_time$dtw) + 50, min(df_time$dtw)) +
    labs(x = 'Year',
         y = 'DTW (ft)') + 
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1))
  plot = ggplotly(gg, tooltip = c('x', 'y', 'wellid')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot
  
}

# PLOTS WELL DTW TIME SERIES FROM DATAFRAME
plotMultipleWells = function(df_time) {
  font = list(
    family = 'Courier',
    size = 15,
    color = 'white')
   label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font)

  gg = ggplot(data = df_time, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    ylim(max(df_time$dtw) + 50, 0) +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well') +  
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1))
  plot = ggplotly(gg, tooltip = c('x', 'y', 'wellid')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot
}

# IDENTICAL FUNCTION TO plotMultipleWells() BUT WILL PLOT ABOVE ZERO ON THE Y-AXIS
plotNegativeWells = function(df_time) {
  font = list(
    family = 'Courier',
    size = 15,
    color = 'white')
  label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font)
  gg = ggplot(data = df_time, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    geom_hline(aes(yintercept = 0), size = 1) +
    scale_y_reverse() +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well') + 
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5), 
          axis.text.x = element_text(color="black", size=14), 
          axis.text.y = element_text(color="black", size=14), 
          axis.title.x = element_text(face="bold", color="black", size=16), 
          axis.title.y = element_text(face="bold", color="black", size=16), 
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1),
          legend.title = element_text(colour="black", size=16, face="bold"),
          legend.text = element_text(colour="black", size=10, face="bold"))
  plot = ggplotly(gg, tooltip = c('x', 'y', 'wellid')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot
}

# RETURNS A TIME SERIES PLOT OF WELLS WITHIN A BUFFER DISTANCE (km)
plotBuffer = function(df, id, buffer) {
  font = list(
    family = 'Courier',
    size = 15,
    color = 'white')
  label = list(
    bgcolor = '#232F34',
    bordercolor = 'transparent',
    font = font)
  well = df %>% filter(wellid == !!id)
  buff = st_buffer(df[well, ], buffer)
  nearby =st_intersection(df, buff)         
  df_time = join_time %>% filter(time_span > 50, wellid %in% nearby$wellid) 
  
  gg = ggplot(data = df_time, aes(x = date, y = dtw)) +
    geom_line(aes(y = dtw, col = wellid), size = 1) +
    ylim(1000, 0) +
    labs(x = 'Year',
         y = 'DTW (ft)',
         col = 'Well') +
    theme_bw() +
    theme(plot.title = element_text(face = 'bold',color = 'black', size = 18, hjust = 0.5),
          axis.text.x = element_text(color="black", size=14),
          axis.text.y = element_text(color="black", size=14),
          axis.title.x = element_text(face="bold", color="black", size=16),
          axis.title.y = element_text(face="bold", color="black", size=16),
          panel.grid.major = element_line(colour = "#808080"),
          panel.grid.minor = element_line(colour = "#808080", size = 1))
  plot = ggplotly(gg, tooltip = c('x', 'y', 'wellid')) %>%
    style(hoverlabel = label) %>% 
    layout(font = font, 
           yaxis = list(fixedrange = TRUE)) %>% 
    config(displayModeBar = FALSE)
  plot
}


# PLOTS A BUFFER AROUND SPECIFIED WELL AND LOCATES OTHER WELLS INSIDE THE BUFFER AREA
buffer_fun = function(df, well, buff, state) {
  buffer = st_buffer(df[well,], buff)
  near1 = st_intersection(df[,], buffer) %>% filter(measurement_dist >= 10)
  
  plot = ggplot() + 
    geom_sf(data = state) +
    geom_sf(data = buffer, fill = NA) + 
    geom_sf(data = near1, col = "red", size = .5) + 
    labs(caption = paste(nrow(near1), 'wells')) +
    theme_void() +
    theme(plot.caption = element_text(size = 22, face = "bold", hjust = 0.5))
  print(plot)
  return(near1)
}

# RETURNS DATAFRAME OF WELLS IN DTW RANGE (MIN - MAX)
dtw_range = function(df, min, max) {
  df = df %>% filter(dtw <= max, dtw >= min) %>% 
    mutate(sd = sd(dtw)) %>% 
    arrange(desc(date))
  return(df)
}

```


























